%% \documentclass{report}
%% 
%% \input{abdou.tex}
%% 
\lecday[2025-04-15]

% \begin{document}

\divider
\textbf{Remainder} : \it(Separable spaces)\normalfont
A toplogical space is said to be separable if it 
contains a countable dense subset.
\divider
\begin{example}
$\RR$  equipped with its usual toplogy is separable
since $Q \subset \RR  $ is countable dense subset of
$\RR$, is a countable dense subset of 
$\RR$. More generally, $\RR ^n  $  
is seperable for all $n \in  \NN $  
(consider the subset $Q^{n} $ of $\RR ^n  $)
\end{example}
\divider
\textbf{
Generalization
}
Every finite dimensional N.V.S (Over $\RR$ or $\CC$  ) 
is separable, since, 
\[
E  \simeq  \KK^n  \simeq  \RR ^n  \simeq  \CC ^n 
\]
\divider
\begin{center}
	\textbf{An important exmaple}
\end{center}
\begin{theorem}[The weirstrass approximation theorem]
	Let $a,b \in \RR  $  with $a < b $, 
	then for every real valued continuous function on 
	$[a,b]  $, 
	there exist a real polynomial sequence $(P_{n}) _{n \in \NN} $  
	which uniformally converges to $f $ on $[a,b]$, in other 
	words, for every $ \veps  > 0 $, there exist a real polynomial
	 $P $ such that 
	 \[
	  \left| f(x) - p(x)  \right| <  \veps  
	  \quad  \left(  \forall  x \in  [a,b]  \right) 
	 \]
	 If $[a,b] = [0,1]$, we can take 
	 \[
	 P_{n}(x)  = 
	 \sum_{k = 0 }^{n}  
	 f(k/n) \binom{n}{k} x^{k}(1-x) ^{n-k}  
	 \]
	 The Bernestein polynomials associated to $f$ 
\end{theorem}
\it Consequence : \normalfont 
let $a,b \in  \RR  $, with $a < b$,
then N.V.S $(\mathcal{C} ^{0}([a,b], \RR , \| . \| _{\infty }))  $, 
is separable. Indeed, the subset of polynomail functions
with rational coefficients on $[a,b]  $ is countable and 
desne in $\left( \mathcal{C} ^{0}([a,b], \RR ), 
\| . \|_{\infty } \right) $  
\begin{definition}[]
Let $E $  be a N.V.S over $\KK=\RR  $  or $\CC $ 
\begin{enumerate}
\item A subset $S $ of $E $ is said to be total if 
	its span (i.e., the set of finite linear 
	combinations of elements of $E $ ) is dense.
\item A Hamel basis of $E $ is linearly independent 
	subset of $E $ which spans $E $ 
	(The concept already known in Linear Algebra-Algebra2)
It follows from Zorn's lemma that every vector space has 
a Hamel basis and that two Hamel bases of a same vector space
are necessarily equinumerous.
\item A schauder basis of $E $ is a sequence 
	$(l_{n}) _{n \in \NN} $  of $E $ such that 
	for each vector $x \in E $, there exists 
	a unique sequence $(\lm_{n}) _{n \in \NN} $  
	of scalars such that 
	\[
	x = \sum_{n=0}^{\infty}  \lm_{n} l_{n}
	\]
	that is, 
	\[
	\| x- \sum_{n=1}^{N} l_{n} \lm_{n} \| \rightarrow 0 
	\quad 0 \text{ as } N \rightarrow  \infty 
	\]
\end{enumerate}
\end{definition}
\begin{center}
	\textbf{Remark : }
\end{center}
\begin{enumerate}
\item Its easy to show that if a N.V.S $E $ has a Schauder basis
	then its separable (Exercise)
\item A Hamel basis (if its finite or countable) of a N.V.S 
	is always Schauder basis (obvious) but the converse is 
	false (see below!)
\item In a finite dimensional N.V.S the concept of Hamel basis and 
	Schauder basis coincides
\end{enumerate}
\begin{example}
\begin{enumerate}
\item (In relation with Fourier series let $p > 1 $, 
	It's show showed that the trigonometric, 
	\[
	1, \cos{(x) }, \sin{(x) }, \hdots   
	\]
	is a Schauder basis of the $\RR$-N.V.S $ L^{p}([0, 2\pi]) $, 
	\[
		L^{p}([0, 2\pi])  = 
		\left\{ f : [0, 2\pi] \rightarrow  \RR 
		\text{ s.t. } 
	\int_{0}^{2\pi} \left| f(x) \right|^{p} d(\mu (x) )  
	< \infty 
\right\} 
	\]
	with the norm $\| . \|_{p} $)
\item   Let $C_{0} $  denote the 
	$\RR  $-vector space of real sequences
	which converge to $0$ and let 
	\[
	\begin{array}{cccc}
	      \| . \| _{\infty } : &  
				   C_0& \longrightarrow & 
				   [0,\infty ]\\
	
	           &  x=(x_{n}) _{n \in \NN}  & \longmapsto     & 
		\| x \| _{\infty } := 
		\sup_{n \in  \NN} \left| x_{n} \right|\\ 
	\end{array}
	\]
\end{enumerate}
It's obvious that $\| . \| _{\infty } $  
is a norm on $C_0 $ (In fact $C_0 $ is a normed subspace of 
$(l^{\infty }, \| . \| _{\infty })  $), where,
\[
	l^{\infty } = 
	\left\{ \text{ the real bounded sequences }  \right\}
\]
for all $n \in  \NN $, let,
\[
l^{(n) } = (l_{i}^{(n) })  _{i \in  \NN}
\]
be the real sequence of $C_0 $ defined by, 
\[
l_{i}^{(n) } := 
\left\{ 
	1 
	\\
	0
\right\}
\begin{gathered}  
i = n \\  
0 \text{ else } 
\end{gathered} = 
(0, 0, \hdots , 0, 0, \ddots ) 
\in  C_0
\]
Its clear that $(e^{(n) })_{n \in \NN}  $  is linearly independent 
and is not a Hamel basis of $C_0$. Because 
\[
	\left\langle e^{(n) }, n \in \NN \right\rangle  = 
	C_{00} \neq C_0 
\]
where 
\[
C_{00} = 
\left\{ \text{ real sequences $(u_{N}) _{n \in \NN} $, for 
$u_{n}=  0 $ for $n $  sufficiently large }  \right\}
\]
$C_{00} \neq  C_0 $  
since we have for example 
$(\frac{1}{n}) _{n \in  \NN} \in C_0 $, but 
$(\frac{1}{n})_{n \in \NN} \notin  C_{00}$. 

Next, for any $x = (x_{n}) _{n \in \NN}  \in  C_0$, 
we have for $n \in \NN $,
\begin{align*}
	\| x - \sum_{n=1}^{N}  x_{n}e^{(n) } \| _{\infty } &=
	\| (x_1, x_2, \hdots ) - 
	(x_1, \hdots , x_{N}, 0, \hdots ) \|_{\infty }    \\
	&=  \| (0, \hdots , 0, x_{N+1}, \hdots )  \| _{\infty } \\
	&= \sup_{n \geq N+1} \left| x_{n} \right|    
\end{align*}
hence,
\begin{align*}
	\lim_{n \to \infty } 
	\| x - \sum_{n=1}^{N} x_{n}e^{(n) } \|  &= \lim_{n  \geq N+1}  
\sup_{}  \left| x_{n} \right| \\
						&=
						\overline{
						\lim_{n \to \infty} } 
						\left| x_{n} \right| \\ &= \lim_{n \to \infty } 
	\left| x_{n} \right| = 0
\end{align*}
This implies that the sequence 
$\left( \sum_{n=1}^{N} x_{n}e^{(n) } \right)_{n \in \NN} $   
of $C_0 $  is convergent to $x$. 

Equivalently, the series $\sum_{n=0}^{\infty}  x_{n}e^{(n) } $  
of $E $ is convergent to $x $, i.e. 
\[
x = \sum_{n=0}^{\infty}  
x_{n}e^{(n) } \quad \text{(in $C_0 $ )} 
\]
Let us show the uniqueness of a such representation of 
$x \in  C_0 $. Suppose that $x \in  C_0 $  is representable
as 
\[
x = \sum_{n=0}^{\infty}  \al_{n} e^{(n) } = 
\sum_{n=0}^{\infty}  \bb_{n} e^{(n) } \quad 
\left( \al_{n}, \bb _n \in  \RR, \forall n \in  \NN\right)
\]
we have for $n \in \NN $, 
\begin{align*}
\| \sum_{i=1}^{N} \al_{i} e^{(i) } - \sum_{i=1}^{N} \bb_{i} e^{(i) }\|   \\
&= \| \sum_{i=1}^{N} \left( \al_{i} - \bb_{i} \right)e^{i} \|  = 
\max_{ 1 \leq i \leq N} 
\left| \al_{i} - \bb_{i} \right|
\end{align*}                     
So for all $n, N \in \NN $, with $n \leq N $, we have, 
\[
\left| \al_{n} - \bb_{n} \right| \leq 
\max _{1 \leq i \leq N} 
\left| \al_{i} - b_{i} \right| =
\| \sum_{i=1}^{N} \al_{i} e^{(i)} - \sum_{i=1}^{N} \bb_{i} e^{(i) } \|  _{\infty }
\text{ By taking } N \rightarrow \infty 
\]
we get that, 
$\left| \al_{n} - \bb_{n} \right|  \leq  0 $, 
thus we have that,
\[
	\al_{n} = \bb_{n} \quad \quad \left( \forall n \in \NN \right)
\]
Thus, the representation of $x$, 
$\sum_{n=1}^{\infty } x_{n} e^{(n) } $ is unique. 

Consequently, $(e^{(n) })_{n \in \NN}  $  is a Schauder basis
of $C_0$ 
\end{example}
\divider
\chapter{Fundamental Theorems on banach spaces : }
\begin{itemize}
	\item The open mapping theorem.
	\item The closed graph theorem.
	\item The Banach-SteinHauns Theorem
	\item The Hahn-Banach 
\end{itemize}
\section{The open mapping theorem}
\textbf{Reminders : } 
A mapping $f $ from a toplogical space $X $ into a toplogical
space $Y $ is said to be an open mapping. if the image by $f $
of every open subset of $X $ is an open subset of $Y $ 
\begin{theorem}[(The open mapping theorem-Schaunder]
	Let $f $ be a continious linear mapping from 
	a Banach space $E $ to a Banach space $F $. Then
	the two following properties are equivalent,
	\begin{enumerate}[i]
	\item  $f $ is surjective
	\item $f $ is an open mapping 
	\end{enumerate}
\end{theorem}
\begin{proof}
\begin{center}
	$(ii)  \implies  (i) $  
\end{center}
We argue by contradiction. Suppose that $f $ is an open mapping
that $f $ is not surjective $(\text{ i.e. } . f(E)  \neq  F)  $, 
so $f(E) $, is a proper subspace of $F $, implying (see
the tutorial worksheet number $1 $ ), that 
\[
	int(f(E) )  = \emptyset 
\]
On the other hand, 
since $f $ is an open mapping and $E $ is open
in $E $ then $f(E)  $  is open in $F $, thus 
$int ( f(E))  = f(E)  $, Hence $f(E) = \emptyset   $, which is a contradiction.

\begin{center}
	$(i)  \implies  (ii)  $ 
\end{center}
we need preliminarry results.


\begin{theorem}[]
Let $E $ and $F $ be two N.V.S and 
$ f : E \longrightarrow F $ be a linear mapping 
then the two following properties are equivalent,
\begin{enumerate}[i]
\item  $f $ is an open mapping 
\item $\exists r > 0 $ such that 
	\[
	B_{F}(0_{F},r) \subset   f\left( B_{E}(0_{E},1)  \right) 
	\]
\end{enumerate}
\end{theorem}
\begin{proof}
	\begin{center}
		$(i)  \implies (ii)  $  
	\end{center}
	Suppose that $f $ is an open mapping. Since
	$B_{E}(0_{E}, 1)  $  is an open 
	subset of $E $, then 
	$f(B_{E}(0_{E},1) )  $  is an open subset of 
	$F $.
	So since,
	\[
	0_{F} = f(0_{E})  \in  f(B_{E}(0_{E},1) ) 
	\]
         then $f(B_{E}(0_{E},1) )  $  is a neighborhood of $0_{F} $, 
	 that is $\exists r > 0 $ such that 
	 \[
	 B_{F}(0_{F},r)   \subset f(B_{E}(0_{E},1) )  
	 \]
	 as required.
\end{proof}
\end{proof}
% \end{document}
